Reinforcement Learning Report

### 2022-08-07 report/rl_qf_report.tex
update training result, finish Reinforce algorithm, working on PPO algorithm

\documentclass[11pt]{article}
\usepackage{amsmath, amsfonts, bm, listings, enumitem, geometry, titlesec,
array, makecell, caption, graphicx, float, algorithm, algpseudocode}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta}
\usepackage{hyperref}
\geometry{margin=0.8in}
\graphicspath{ {./imgs/} }
%\renewcommand{\familydefault}{\sfdefault}

\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1.3cm, 
text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>={Stealth[scale=1.1]}]

@@

\section{RL with Actor-Critic}
In this project, we consider the episodic learning procedure, where each 
episode has finite time steps, \(N\), to be specified as a constant, and we 
will also reset the environment at the start of each episode. The algorithm 
below is to use RL, with actor-critic method, and with the help of what we 
have built in previous sections, to solve the problem:
\begin{algorithm}[H]
\caption{RL-Actor-Critic-for-financial-problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an \(x\) (first customer of the episode)
    \State Initialize \(S\) (first state of the episode)
    \Loop \hspace{0.3mm} (for each time step \(i\), \(N\) in total)
        \State \(c\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(c\), observe \(S'\), \(R\) from environment
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{i}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}
\noindent The following graph illustrates the algorithm:
\vspace{5mm}
\begin{center}
    \begin{tikzpicture}[node distance=2cm]
        \node (agent) [process, minimum width=6cm, minimum height=3cm,
            label={[anchor=north east,inner sep=7pt]north east:Agent}]{};
        \node (pi) [process, left of=agent, xshift=0.5cm, yshift=-0.3cm] 
            {\(\pi\left(c\middle|s,\bm{\theta}\right)\)};
        \node (v) [process, right of=pi, xshift=1cm] 
            {\(\hat{v}\left(s,\bm{w}\right)\)};
        \node (env) [process, below of=pi, yshift=-2.5cm] {Environment};

@@

>>>>

\documentclass[11pt]{article}
\usepackage{amsmath, amsfonts, bm, listings, enumitem, geometry, titlesec,
array, makecell, caption, graphicx, float, algorithm, algpseudocode, siunitx}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta}
\usepackage{hyperref}
\geometry{margin=0.8in}
\graphicspath{ {../data/} }

%\renewcommand{\familydefault}{\sfdefault}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1.3cm, 
text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>={Stealth[scale=1.1]}]

@@

\section{RL with Actor-Critic}
\subsection{The Algorithm}
In this project, we consider the episodic learning procedure, where each 
episode has finite time steps, \(T\), to be specified as a constant, and we 
will also reset the environment at the start of each episode. The algorithm 
below is to use RL, with actor-critic method, and with the help of what we 
have built in previous sections, to solve the problem:
\begin{algorithm}[H]
\caption{Actor-Critic for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an \(x\) (first customer of the episode)
    \State Initialize \(S\) (first state of the episode)
    \Loop \hspace{0.3mm} (for each time step \(t\), \(T\) in total)
        \State \(c\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(c\), observe \(S'\), \(R\) from environment
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{t}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}
\noindent The following graph illustrates the algorithm:
\vspace{5mm}
\begin{center}
    \begin{tikzpicture}[node distance=2cm]
        \node (agent) [process, minimum width=6cm, minimum height=3cm,
            label={[anchor=north east,inner sep=7pt]north east:Agent}]{};
        \node (pi) [process, left of=agent, xshift=0.5cm, yshift=-0.3cm] 
            {\(\pi\left(c\middle|s,\bm{\theta}\right)\)};
        \node (v) [process, right of=pi, xshift=1cm] 
            {\(\hat{v}\left(s,\bm{w}\right)\)};
        \node (env) [process, below of=pi, yshift=-2.5cm] {Environment};

@@

\subsection{Training Result}
The following graph shows the training result of the above algorithm (300
moving average reward vs iteration), with 600 iterations (5 episodes per 
iteration), \(T=300\), \(\gamma=0.99\), 
\(\alpha_\theta=\alpha_{\bm{w}}=\num{3e-4}\).
\begin{center}
    \includegraphics[scale=0.67]{a2c_out}
\end{center}
The graph shows that the moving average reward converges within only 50 
iterations, that is, 250 episode, 75000 parameter updates for each network, 
and stays quite stable afterwards.

\section{Actor-Critic, Reinforce, and PPO}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms.
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, A_0, R_1, ..., S_{T-1}, A_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \Loop \hspace{0.3mm} (for each time step \(t\), \(T\) in total)
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(A_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{\(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(c\middle|s, \bm{\theta}\right)\)
    \State Compute discounted returns \(G_k\), the same way as in Reinforce
    algorithm
\EndFor
\end{algorithmic}
\end{algorithm}
